---
title: "p8105_hw3_yz4720"
author: "Jasmine Zhang"
date: "2023-10-14"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1
```{r}
data(instacart)
```
The `instacart` dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, and each row represents an item from an order. The variables include order id, product id, product name, number identifier of the item in the order, whether the item is a reorder, and user id. There is also information on the day and time of the order, and number of days since prior order, location of the aisle and department with their corresponding id. In total, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```
There are 134 aisles in total. Most items are ordered from fresh vegetables and fresh fruits aisles.

```{r}
instacart |> 
  count(aisle) |> 
  filter(n>10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |>  
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
  
```

```{r}
instacart |> 
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```

